---
title: "TCGA HNSC screen method"
format: pdf
editor: visual
---

# Using screen methods to select miRNA \~ survival data

```{r echo=F}
######################################
library(MASS)
library(splines)
library(glmnet)
library(Matrix)
library(mvtnorm)
library(survival)
library(glasso)
library(network)
library(ncvreg)
library(Hmisc)
library(huge)
library(corrplot)
library(GEInter)
library(SIS)
########################################
```

# Set up the data

```{r warning=F,echo=F}
##Loading data

#TCGA<-read.csv("df258_sc.csv", header = TRUE) ##1. according to comparing case and control to selected miRNA(important***)
TCGA<-read.csv("df290_sc.csv", header = TRUE) ## 2. only clearn the NA value of miRNA





# Exclude the first column
TCGA <- TCGA[,-1]
#TCGA <- TCGA[,-2]

library(survival)

 #Create a survival object with time and event variables
my.survival <- Surv(TCGA$OS.time, TCGA$OS)

#Fit the Kaplan-Meier survival curve
my.km <- survfit(my.survival ~ 1)

# Plot the survival curve
plot(my.km)

##checking data
dim(TCGA) ## 326 row, 260-3 number of genes

length(TCGA[,3])

##Transferring unit, from day to months
TCGA[,3]=TCGA[,3]/30+runif(dim(TCGA)[1],1e-05,1e-03) #months

##creating the plot for survival time (months) of each subject
par(mfcol=c(1,2))
subject=c(1:dim(TCGA)[1])
plot(subject,TCGA[,3],xlab="subject",ylab="survival time (months)")


##Creating the probability density function
sd=sd(TCGA[,3])
mean=mean(TCGA[,3])
sd3=3*sd+mean

abline(h=sd3,col="red",lwd=2) ## line of standard variance
abline(h=mean,col="blue",lwd=2)## line of mean
d<-density(as.matrix(TCGA[,3]))## denstity function
plot(d,xlab="survival time (months)",ylab="density",main = "TCGA") ## plot of density
curve(dnorm(x,mean,sd),add=T,col="red3",lwd=2) ##"predicted curve"

allp=ncol(TCGA)-3 # number of gene (since the first 3 columns of the original data is not gene)

SS=TCGA[,2]; XX=TCGA[,3] #  SS = os (live or dead)   XX = os.time (how long has the subject alive?),



APL=c(rep(0,allp)) # creating a empty list for 257 genes


## coxph model for every gene
for (j in 1:allp){ 
  #print(j)        
  tempz=(TCGA[,3+j]) ##here is the matrix of gene
  fit=try(coxph(Surv(XX,SS)~tempz)) ## skip error  because sometime there is no result. N/A
  

  APL[j]=summary(fit)$coefficients[,5]    ### taking p-value
}
APL[which(is.na(APL))]=1 ## set N/A = p-value is 1



numsel=ncol(TCGA)-3# number of genes

## order the result (p-value) from 'coxph' model (from the lowest to highest)
ssindex=sort(order(APL)[1:numsel]) 

sindex<-ssindex+3 # number of index  in original data (from the third col to start count)
NTCGA<-cbind(XX,SS,TCGA[,sindex])#time,delta,data (tis is the ordered X matrix)

dim(NTCGA)

tempn<-colnames(NTCGA)
class(tempn)

NN=dim(NTCGA)[1]
sum(!duplicated(NTCGA[,1]))==NN ##check the number matched or not



## include all the subject
trainindex=sort(sample(sample(sample(c(1:NN))))[1:(round(NN*1))])  ## if we change the index of NN*index, we could chose how my percent subject, we want to use.


## test index.........?????? what is this mean?    
testindex=c(1:NN)[-trainindex]## the orgin code have the testing part. I think we could leave it.

    
xo<-as.matrix(NTCGA[trainindex,-c(1:2)])##author created this matrix just for train matrix.
dim(xo)
  
    time=NTCGA[trainindex,1]; delta=NTCGA[trainindex,2] ## "1" is for time, '2' is for status.
    
    yy<-NTCGA[trainindex,1:2]; colnames(yy)=c("time","status")
    
    #cen.r[kkk]=1-mean(delta) ##????? what is this for?
    
    n = dim(xo)[1]
    p = dim(xo)[2]
    
PASg=function(z) {
      g=paste("G", z, sep = "")
      return(g)
    }
    newxo<-cbind(yy,xo) ## yy survial dta and xo is gene matrix
    

newxo<-as.data.frame(newxo)
colnames(newxo)<-c("time","event",PASg(c(1:p)))

```

# 8 screen methods

##### marginal partial likelihood screening (mPL)

```{r warning=F, echo=T}
mPL=numeric(p)

    for(j in 1:p){
      
      ###marginal partial likelihood screening
      fit1 = try(coxph(Surv(time,delta)~xo[,j]))
      if(inherits(fit1,"try-error")){
        mPL[j]= -100000
      } else {
        mPL[j] = fit1$loglik[2] ###log partial likelihood for jth covariate
      }
    }

AZ11=xo[,order(mPL,decreasing=T)[1:numsel]]
dim(AZ11)
# PL  Lbeta1
    fitcv1=ncvsurv(X=AZ11, yy, penalty="MCP", nlambda=100)
    lam1=fitcv1$lambda[which.min(BIC(fitcv1))] #AIC
    Lbeta1=as.matrix(coef(fitcv1, lambda=lam1)) 
    names1=colnames(AZ11)[which(Lbeta1[,1]!=0)]
    Lbeta1[Lbeta1[,1]!=0,1]
    
    summary(fitcv1,lam1)
    
rm(AZ11) 
```

##### SIS

```{r warning=FALSE ,echo=T}
#
sis <- function(df, time, status, features){
  out <- apply(df[,features], 2, function(d) abs(cor(d, df[,time])))
  return(out)
}

## filled the new X matrix into the SIS model
NEWSIS<-sis(newxo,"time","event",colnames(newxo)[-c(1:2)])


##the number of numsel mean that how many gene will be include in this model
AZ22=xo[ ,order(NEWSIS,decreasing=T) [1:numsel] ]

##SIS
 fitcv2=ncvsurv(X=AZ22, yy, penalty="MCP", nlambda=100)
  lam2=fitcv2$lambda[which.min(BIC(fitcv2))] #AIC
    Lbeta2=as.matrix(coef(fitcv2, lambda=lam2))
    names2=colnames(AZ22)[which(Lbeta2[,1]!=0)]
    Lbeta2[Lbeta2[,1]!=0,1]
    
    summary(fitcv2,lam2)

rm(AZ22)  
```

##### FAST

```{r warning=FALSE, echo=T}
#FAST function
FAST <- function(df, time, status, features){
  uniahaz <- ahaz::ahaz(surv = Surv(df[,time],df[,status]), X = df[,features], univariate = TRUE)
  out     <- abs(uniahaz$d)
  return(out)
}

NEWFAST<-FAST(newxo,"time","event",colnames(newxo)[-c(1:2)])

AZ33=xo[,order(NEWFAST,decreasing=T)[1:numsel]]

 # FAST
   fitcv3=ncvsurv(X=AZ33, yy, penalty="MCP", nlambda=100)
    lam3=fitcv3$lambda[which.min(BIC(fitcv3))] #AIC
   Lbeta3=as.matrix(coef(fitcv3, lambda=lam3))
   names3=colnames(AZ33)[which(Lbeta3[,1]!=0)]
   Lbeta3[Lbeta3[,1]!=0,1]
   
   
   summary(fitcv3, lam3)
    
rm(AZ33)
```

##### Method of RCDCS

```{r warning=FALSE,echo=T}
# Chen, Chen & Wang, CSDA 2018
RCDCS <- function(df, time, status, features){
  df     <- df[order(df[,time]),]
  # Fn(t)
  kme    <- survfit(formula(paste("Surv(",time,",",status,")~1")), df)
  df$FnT <- 1 - summary(kme, censored=T, times=df[,time])$surv
  # replace xj by Fjn(x)
  df[,features] <- apply(df[,features], 2, function(d) ecdf(d)(d))
  # distance correlation w/o asymptotic results
  out <- apply(df[,features], 2, function(d) Rfast::dcor(as.matrix(d), df$FnT)$dcor^2) # ^2(?)
  return(out)
}

NEWRCDCS<-RCDCS(newxo,"time","event",colnames(newxo)[-c(1:2)])


##the number of numsel mean that how many gene will be include in this model
AZRCDCS=xo[,order(NEWRCDCS,decreasing=T)[1:numsel]] 
dim(AZRCDCS)# included all the information of data

     #RCDCS
    fitcv4=ncvsurv(X=AZRCDCS, yy, penalty="MCP", nlambda=100)
    lam4=fitcv4$lambda[which.min(BIC(fitcv4))] #AIC
    Lbeta4=as.matrix(coef(fitcv4, lambda=lam4))
    names4=colnames(AZRCDCS)[which(Lbeta4[,1]!=0)]
    
    Lbeta4[Lbeta4[,1]!=0,1]
    summary(fitcv4,lam4)
    
 

```

##### CRCDCS

```{r warning=FALSE, echo=T}
# Chen, Chen & Wang, CSDA 2018
CRCDCS <- function(df, time, status, features, q=seq(0.1,0.9,0.05)){
  df     <- df[order(df[,time]),]
  # Fn(t)
  kme    <- survfit(formula(paste("Surv(",time,",",status,")~1")), df)
  df$FnT <- 1 - summary(kme, censored=T, times=df[,time])$surv
  # Qt(t)
  qt     <- quantile(kme,q)$quantile # probs apply to F(t)=1-S(t)
  # nonmissing sample distribution estimates
  q      <- q[!is.na(qt)]
  qt     <- qt[!is.na(qt)]
  # wih
  wmat <- matrix(NA, ncol=length(q),nrow=nrow(df))
  for (i in 1:length(q)) {
    wh       <- ifelse(df[,status]==1 | df$FnT > q[i], 1, (q[i] - df$FnT)/(1 - df$FnT))
    wmat[,i] <- q[i] - wh * as.numeric(df[,time] <= qt[i])
  }
  # replace xj by Fjn(x)
  df[,features] <- apply(df[,features], 2, function(d) ecdf(d)(d))
  # distance correlation
  out <- apply(df[,features], 2, function(d) Rfast::dcor(as.matrix(d), wmat)$dcor^2) # ^2(?)
  return(out)
}


NEWCRCDCS<-CRCDCS(newxo,"time","event",colnames(newxo)[-c(1:2)])


AZ55=xo[,order(NEWCRCDCS,decreasing=T)[1:numsel]]


fitcv5=ncvsurv(X=AZ55, yy, penalty="MCP", nlambda=100)
   lam5=fitcv5$lambda[which.min(BIC(fitcv5))] #AIC
    Lbeta5=as.matrix(coef(fitcv5, lambda=lam5))
    names5=colnames(AZ55)[which(Lbeta5[,1]!=0)]
    Lbeta5[Lbeta5[,1]!=0,1]
    
    summary(fitcv5,lam5)

```

##### CINDEX

```{r warning=FALSE, echo=T}
#
CoxConc <- function(df, time, status, features){
  out <- apply(df[,features], 2, function(d) summary(coxph(Surv(df[,time], df[,status]) ~ d))$concordance[1])
  return(out)
}

NEWCINDEX<-CoxConc(newxo,"time","event",colnames(newxo)[-c(1:2)])

AZ66=xo[,order(NEWCINDEX,decreasing=T)[1:numsel]]

 fitcv6=ncvsurv(X=AZ66, yy, penalty="MCP", nlambda=100)
    lam6=fitcv6$lambda[which.min(BIC(fitcv6))] #AIC
   Lbeta6=as.matrix(coef(fitcv6, lambda=lam6))
   names6=colnames(AZ66)[which(Lbeta6[,1]!=0)]
   Lbeta6[Lbeta6[,1]!=0,1]
    
   summary(fitcv6,lam6)
```

##### mIPCW_tau

```{r warning=FALSE, echo = T}
mIPCW_tau = numeric(p) 

 ##################################################################
    ###compute Kanplan-Meier estimate for censoring distribution
    ##################################################################
    km=survfit(Surv(time,1-delta)~1)
    kmT=summary(km)$time ; kmS=summary(km)$surv
    kms=c(rep(0,n))
    
    for (i in 1:n){
      temp=sum(time[i]>=kmT)
      if (temp==0) kms[i]=1 else kms[i]=kmS[temp]
    }

     ############################################################
      #1, > >
      w=c(delta/kms^2); w[is.infinite(w)]=0; w[is.na(w)]=0
      v=c(time)
      W=matrix(c(w),n,n)
      X=matrix(c(xo[,j]),n,n)
      V=matrix(c(v),n,n)
      XX=sign(t(X)-X)
      VV=sign(t(V)-V)
      
      f=n^2-n
      XX[XX==-1]=0
      VV[VV==-1]=0
      A1W=sum(W*XX*VV)/f
      A1=sum(XX*VV)/f
      
      #3, < >
      XX=sign(t(X)-X)
      VV=sign(t(V)-V)
      XX[XX==1]=0; XX[XX==-1]=1
      VV[VV==-1]=0
      B1W=sum(W*XX*VV)/f
      B1=sum(XX*VV)/f
      
      ########################################################################
      mIPCW_tau[j] = abs(2*A1W-2*B1W) #### modified

 
      
      AZ77=xo[,order(mIPCW_tau,decreasing=T)[1:numsel]]

   # IPCW-tau
    fitcv7=ncvsurv(X=AZ77, yy, penalty="MCP", nlambda=100)
    
  
    
    lam7=fitcv7$lambda[which.min(BIC(fitcv7))] #AIC
    
    Lbeta7=as.matrix(coef(fitcv7, lambda=lam7))
    
    names7=colnames(AZ77)[which(Lbeta7[,1]!=0)]
    Lbeta7[Lbeta7[,1]!=0,1]
    
    summary(fitcv7,lam7)
    

    
 



```

##### IPCW-tau(NPN-MB)

```{r warning=FALSE, echo =T}
 #################### GLASSO  #####################################
    #################################################################
    
    # step1
    I=matrix(1,p,1);C=matrix(mIPCW_tau,1,p)
    tempC=I%*%C
    tempCsum=apply(tempC,1,sum)
    CC=tempC/tempCsum
    
    ##################################IPCW-tau(NPN-MB)
    # transform the data using the shrunken ECDF
    #npn.sxo=huge.npn(x=sxo, npn.func="shrinkage")
    # transform the non-Gaussian data using the truncated ECDF
    ##################################################################
    
    sxo=scale(xo)
    npn.sxo=huge.npn(x=sxo, npn.func="truncation")
    
    out.npn.mb=huge(x=npn.sxo,nlambda=30,lambda.min.ratio=0.1,method="mb", scr=TRUE)
    #plot(out.npn.mb)
    #huge.roc(out.npn.mb$path,L$theta)
    
    npn.mb.ric=huge.select(out.npn.mb, criterion="ric")
    #plot(npn.mb.ric)
    #print(npn.mb.ric$opt.lambda)
    
    wpc2=npn.mb.ric$refit+1-1; diag(wpc2)=1
    g=network(wpc2)
    plot(g,main="Estimated NPN-MB network")
    
    pcsum2=apply(wpc2,1,sum)
    A2=wpc2/pcsum2
    W=seq(0,1,1/100)
    VV=matrix(0,length(W),p)
    
    
    for (k in 1:length(W)){
      S=as.matrix((1-W[k])*CC+W[k]*A2)
      V=matrix(1/p,1,p)
      for (i in 1:100){
        V=V%*%S
        V=V/sum(V)
      }
      VV[k,]=V
    }
    
    
    FV2=apply(VV,2,mean)
    


 AZ88=xo[, order(FV2,decreasing=T)[1:numsel]]

# NPN-MB
    fitcv8=ncvsurv(X=AZ88, yy, penalty="MCP", nlambda=100)
    
    
    lam8=fitcv8$lambda[which.min(BIC(fitcv8))] #AIC
    Lbeta8=as.matrix(coef(fitcv8, lambda=lam8))
    names8=colnames(AZ88)[which(Lbeta8[,1]!=0)]
    Lbeta8[Lbeta8[,1]!=0,1]
    
    summary(fitcv8,lam8)
    

    

```

# Using SIS to select miRNA \~ LVI

##### Using SIS function

```{r warning=FALSE ,echo =T}
y=Surv(newxo$time,newxo$event)
dim(y)
x=newxo[,-c(1:2)]
x<-as.matrix(x)

sec_model<-tune.fit(x,y,family = 'cox',penalty = 'MCP',tune='bic')
#sec_model<-SIS(x,y,family = 'cox',penalty = 'MCP',tune='bic')

fitcv0<-sec_model$fit
lamb0<-fitcv0$lambda

lam0=fitcv0$lambda[which.min(BIC(fitcv0))]#AIC
summary(fitcv0,lam0)

#colnames(TCGA[188+3])
#colnames(TCGA[195+3])

```

##### Without using SIS function

```{r warning=FALSE ,echo =T}
sis <- function(df, time, status, features){
  out <- apply(df[,features], 2, function(d) abs(cor(d, df[,time])))
  return(out)
}
## filled the new X matrix into the SIS model
NEWSIS<-sis(newxo,"time","event",colnames(newxo)[-c(1:2)])
##the number of numsel mean that how many gene will be include in this model
AZ22=xo[ ,order(NEWSIS,decreasing=T) [1:numsel] ]
##SIS
 fitcv2=ncvsurv(X=AZ22, yy, penalty="MCP", nlambda=100)
  lam2=fitcv2$lambda[which.min(BIC(fitcv2))] #AIC
    Lbeta2=as.matrix(coef(fitcv2, lambda=lam2))
    names2=colnames(AZ22)[which(Lbeta2[,1]!=0)]
    Lbeta2[Lbeta2[,1]!=0,1]
    summary(fitcv2,lam2)
 
    
       
```

# Looking for miRNA assocation with lymphovascular invasion status?

##### loading data set with lymo variable

```{r warning=FALSE, echo =T}
##Loading data sets

##1. according to comparing case and control to selected miRNA(important***)this is fine
#lymp <-read.csv("df258_LVI.csv", header = TRUE) 

## 2. only clearn the NA value of miRNA
lymp<-read.csv("df290_LVI.csv", header = TRUE) 
```

##### using SIS function to apply lymp data (checking the result)

```{r warning=FALSE ,echo =T}

# Exclude the first column
lymp <- lymp[,-1]

##check size
dim(lymp) 

##creat lymp data set
lymp<-lymp[,-c(3:4)]
dim(lymp) # 326 subject & 259 (292) variables (the number of gene should be 292-2 = 290)


x<-as.matrix(lymp[-c(1:2)])
dim(x)
y<-lymp$lym_invasion


sis_result <- SIS::SIS(x, y, family = "binomial",penalty = "MCP",tune = "bic")

fitcv3<-sis_result$fit
lamb3<-fitcv3$lambda
lam3=fitcv3$lambda[which.min(BIC(fitcv3))]#AIC

summary(fitcv3,lam3)
```

##### using tune.fit function to apply lymp data (checking the result)

```{r warning=FALSE ,echo =T}
sis_result02<-tune.fit(x,y,family = 'binomial',penalty = 'MCP',tune='bic')

fitcv04<-sis_result02$fit
lamb04<-fitcv04$lambda

lam04=fitcv04$lambda[which.min(BIC(fitcv04))]#AIC
summary(fitcv04,lam04)
```
